Technical Report

### 2018-03-08 fully-abstract-comp/technical-report.tex
Some work on FTLR proof.
Leave out rwx pointers from the value relation.
Add a lemma to be proven.
Add a TODO about readXCondition

The $\decInstr{}$ function must be surjective and injective for all non-$\tfail$ instructions. The $\encInstr{}$ function must be injective. Further $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
\[
  \decInstr{}(\encInstr{i}) = i
\]

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
\[
  \lrv(H_\sigma,W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{c_S},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
      \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
%
    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    % \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \npair{(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l}
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \npair{(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l }
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
        \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
                                   \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

@@

\[
  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
      \qquad \pwheap(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \stackReadCond{W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\linear, \pwfree} \ldotp \\
      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
      \qquad \pwfree(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\lin, \pwheap}, \ms \ldotp\\
      \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
      \quad \pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}
    \end{array}
  \right\}
\]

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
% Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps. It expresses that any reasonable executable capability is safe to use as the pc capability. 
\dominique{I'm following the example of the previous FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
\begin{theorem}[FTLR]
  \label{thm:ftlr}
  \[
    \begin{array}{l}
      \forall n,W,\lin,\baddr,\eaddr,\aaddr \ldotp \\
      \quad \left(
      \begin{array}{l}
        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}
      \end{array}
      \right)\\
      \quad\Rightarrow 
      \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
                      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)

    \end{array}
  \]
  \dominique{6-3-2018: $\readCond{\lin,W}$ and $\xReadCond{\lin,W}$ seem contradictory.
    I suspect the latter needs to go? I suspect we don't need to assume the context has no way to modify its own code.}
\end{theorem}

>>>>

The $\decInstr{}$ function must be surjective and injective for all non-$\tfail$ instructions.
For any $c \in \Caps$, we have that $\decInstr{c} = \tfail$.
The $\encInstr{}$ function must be injective.
Further $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
\[
  \decInstr{}(\encInstr{i}) = i
\]

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
\[
  \lrv(H_\sigma,W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{c_S},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
      \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
%
    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    % \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \npair{(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l}
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \npair{(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l }
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
        % we are excluding rwx pointers.
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

@@

\[
  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
      \qquad \pwheap(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \stackReadCond{W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\linear, \pwfree} \ldotp \\
      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
      \qquad \pwfree(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\lin, \pwheap}, \ms \ldotp\\
      \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
      \quad \pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}
    \end{array}
  \right\}
\]
\dominique{The above needs to require that $\ms$ contains safe values somehow.}

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
% Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps. It expresses that any reasonable executable capability is safe to use as the pc capability. 
\dominique{I'm following the example of the previous FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
\begin{theorem}[FTLR]
  \label{thm:ftlr}
  \[
    \begin{array}{l}
      \forall n,W,\lin,\baddr,\eaddr,\aaddr \ldotp \\
      \quad \left(
      \begin{array}{l}
        % \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}
      \end{array}
      \right)\\
      \quad\Rightarrow 
      \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
                      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)

    \end{array}
  \]
  \dominique{6-3-2018: I removed $\readCond{\lin,W}$ and replaced it with $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
\end{theorem}
